---
title: "multilevel_models"
author: "Alexander Berry"
date: "4/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(blme)
library(corrplot)
library(BBmisc)
```
Load in data:
```{r}
parent_dir = dirname(getwd())
data_dir = paste(parent_dir, "data", sep="/")
df = read.csv(paste(data_dir, "african_crises.csv", sep="/"))
```
Preprocessing
```{r}
distinct.years = df %>% group_by(year) %>% summarise(n_distinct(country)) 
valid_years = distinct.years %>% filter(.[[2]] > 12) %>% select(year)
df$exch_usd = normalize(df$exch_usd, method="range")
df$inflation_annual_cpi = normalize(df$inflation_annual_cpi, method="range")
df$gdp_weighted_default = normalize(df$gdp_weighted_default, method="range")
pre.df = df %>% inner_join(y = valid_years, by='year') %>% arrange(year)
rownames(pre.df) <- 1:nrow(pre.df)
pre.df = pre.df %>% mutate(banking_crisis = ifelse(banking_crisis == "crisis", 1, 0))
pre.df$currency_crises[pre.df$currency_crises == 2] <- 1

write.csv(pre.df, paste(data_dir, "preprocessed_african_crises.csv", sep="/"))
```
Quick EDA
```{r}
pre.df %>% select(-c("country", "cc3", "case", "year")) %>% cor() -> M
corrplot(M, method="circle")
```
PP II
```{r}
pre.df = transform(pre.df,
                       systemic_crisis = as.factor(systemic_crisis),
                       # year = as.factor(year),
                       country = as.factor(country),
                       banking_crisis = as.factor(banking_crisis),
                       currency_crises = as.factor(currency_crises),
                       inflation_crises = as.factor(inflation_crises),
                       domestic_debt_in_default = as.factor(domestic_debt_in_default),
                       sovereign_external_debt_default = as.factor(sovereign_external_debt_default),
                       independence = as.factor(independence))
```

Systemic Crises
```{r}
systemic_model_all = bglmer(systemic_crisis ~ (1 | country ) + (1 | year) + exch_usd + inflation_annual_cpi + domestic_debt_in_default + gdp_weighted_default + sovereign_external_debt_default + inflation_crises + banking_crisis + currency_crises + independence,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)) )
summary(systemic_model_all)
```


Define multilevel model
```{r}
systemic_model_better = bglmer(systemic_crisis ~ (1 | country ) + (1 | year) + exch_usd + inflation_crises + currency_crises + inflation_annual_cpi + domestic_debt_in_default + sovereign_external_debt_default,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(systemic_model_better)
```
Inflation Crises
```{r}
inflation_model_all = bglmer(inflation_crises ~ (1 | country ) + (1 | year) + exch_usd + inflation_annual_cpi + domestic_debt_in_default + gdp_weighted_default + sovereign_external_debt_default + systemic_crisis + banking_crisis + currency_crises + independence,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)) )
summary(inflation_model_all)
```
Currency Crises
```{r}
currency_model_all = bglmer(currency_crises ~ (1 | country ) + (1 | year) + exch_usd + inflation_annual_cpi + domestic_debt_in_default + gdp_weighted_default + sovereign_external_debt_default + systemic_crisis + banking_crisis + inflation_crises + independence,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)) )
summary(currency_model_all)
```
Banking Crises
```{r}
banking_model_all = bglmer(banking_crisis ~ (1 | country ) + (1 | year) + exch_usd + inflation_annual_cpi + domestic_debt_in_default + gdp_weighted_default + sovereign_external_debt_default + systemic_crisis + currency_crises + inflation_crises + independence,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)) )
summary(banking_model_all)
```
Group countries into regions
```{r}
# country.to.region = function(c.codes){
#   north = c("DZA", "EGY", "MAR", "TUN")
#   center = c("CAF", "CIV", "KEN", "NGA")
#   south = c("AGO", "MUS", "ZAF", "ZMB", "ZWE")
#   regions = vector(mode="character", length=length(c.codes))
#   for(i in length(c.codes)){
#     if(is.element(c.codes[i], north)){
#       regions[i] = "north"
#       cat
#     }
#     else{
#       if(is.element(c.codes[i], center)){
#         regions[i] = "center"
#       }
#       else{
#         if(is.element(c.codes[i], south)){
#           regions[i] = "south"
#         }
#       }
#     }
#   }
#   return(regions)
# }
# country.to.region(pre.df$cc3)
north = c("DZA", "EGY", "MAR", "TUN")
center = c("CAF", "CIV", "KEN", "NGA")
south = c("AGO", "MUS", "ZAF", "ZMB", "ZWE")
pre.df$region[is.element(pre.df$cc3, north)] <- "north"
pre.df$region[is.element(pre.df$cc3, center)] <- "center"
pre.df$region[is.element(pre.df$cc3, south)] <- "south"


pre.df$decade[pre.df$year >= 1950 & pre.df$year < 1960] = "1950s"
pre.df$decade[pre.df$year >= 1960 & pre.df$year < 1970] = "1960s"
pre.df$decade[pre.df$year >= 1970 & pre.df$year < 1980] = "1970s"
pre.df$decade[pre.df$year >= 1980 & pre.df$year < 1990] <- "1990s"
pre.df$decade[pre.df$year >= 1990 & pre.df$year < 2000] <- "1990s"
pre.df$decade[pre.df$year >= 2000 & pre.df$year < 2010] <- "2000s"
pre.df$decade[pre.df$year >= 2010 & pre.df$year < 2020] <- "2010s"
```

Multilevel with regions
```{r}
systemic_model_all_regions = bglmer(systemic_crisis ~ (1 | region ) + exch_usd + inflation_annual_cpi + domestic_debt_in_default + gdp_weighted_default + sovereign_external_debt_default + inflation_crises + banking_crisis + currency_crises + independence,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)) )
summary(systemic_model_all_regions)
```

Multilevel with decades
```{r}
systemic_model_all_decades = bglmer(systemic_crisis ~ (1 | decade ) + exch_usd + inflation_annual_cpi + domestic_debt_in_default + gdp_weighted_default + sovereign_external_debt_default + inflation_crises + banking_crisis + currency_crises + independence,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)) )
summary(systemic_model_all_decades)
```

Both region and decade (all)
```{r}
systemic_model_all_reg.dec = bglmer(systemic_crisis ~ (1 | decade ) + (1 | region ) + exch_usd + inflation_annual_cpi + domestic_debt_in_default + gdp_weighted_default + sovereign_external_debt_default + inflation_crises + banking_crisis + currency_crises + independence,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)) )
summary(systemic_model_all_reg.dec)
```

Better region and decade
```{r}
systemic_model_better_reg.dec = bglmer(systemic_crisis ~ (1 | decade ) + (1 | region ) + exch_usd + inflation_annual_cpi + domestic_debt_in_default + sovereign_external_debt_default + inflation_crises + banking_crisis + currency_crises,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)) )
summary(systemic_model_better_reg.dec)
```

```{r}
systemic_model_best_reg.dec = bglmer(systemic_crisis ~ (1 | decade ) + (1 | region ) + inflation_annual_cpi + domestic_debt_in_default + banking_crisis,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)) )
summary(systemic_model_best_reg.dec)
```
Inflation crisis (region, decade, all)
```{r}
inflation_model_all_reg.dec = bglmer(inflation_crises ~ (1 | decade ) + (1 | region ) + exch_usd + inflation_annual_cpi + domestic_debt_in_default + gdp_weighted_default + sovereign_external_debt_default + systemic_crisis + banking_crisis + currency_crises + independence,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)) )
summary(inflation_model_all_reg.dec)
```
Inflation better
```{r}
inflation_model_better_reg.dec = bglmer(inflation_crises ~ (1 | decade ) + (1 | region ) + exch_usd + inflation_annual_cpi + gdp_weighted_default + sovereign_external_debt_default + systemic_crisis + banking_crisis,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)) )
summary(inflation_model_better_reg.dec)
```
Inflation best
```{r}
inflation_model_best_reg.dec = bglmer(inflation_crises ~ (1 | decade ) + (1 | region ) + exch_usd + inflation_annual_cpi + gdp_weighted_default + sovereign_external_debt_default,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)) )
summary(inflation_model_best_reg.dec)
```

Banking all
```{r}
banking_model_all_reg.dec = bglmer(banking_crisis ~ (1 | decade ) + (1 | region ) + exch_usd + inflation_annual_cpi + domestic_debt_in_default + gdp_weighted_default + sovereign_external_debt_default + systemic_crisis + inflation_crises + currency_crises + independence,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)) )
summary(banking_model_all_reg.dec)
```
Banking better
```{r}
banking_model_better_reg.dec = bglmer(banking_crisis ~ (1 | decade ) + (1 | region ) + exch_usd + inflation_annual_cpi + domestic_debt_in_default + sovereign_external_debt_default + systemic_crisis + inflation_crises + currency_crises ,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)) )
summary(banking_model_better_reg.dec)
```
Banking best
```{r}
banking_model_best_reg.dec = bglmer(banking_crisis ~ (1 | decade ) + (1 | region ) + exch_usd + inflation_annual_cpi + domestic_debt_in_default + sovereign_external_debt_default + systemic_crisis,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)) )
summary(banking_model_best_reg.dec)
```
Currency all
```{r}
currency_model_all_reg.dec = bglmer(currency_crises ~ (1 | decade ) + (1 | region ) + exch_usd + inflation_annual_cpi + domestic_debt_in_default + gdp_weighted_default + sovereign_external_debt_default + systemic_crisis + inflation_crises + banking_crisis + independence,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)) )
summary(currency_model_all_reg.dec)
```
```{r}
currency_model_better_reg.dec = bglmer(currency_crises ~ (1 | decade ) + (1 | region ) + inflation_annual_cpi + domestic_debt_in_default + systemic_crisis + inflation_crises,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(currency_model_better_reg.dec)
```
Currency best
```{r}
currency_model_best_reg.dec = bglmer(currency_crises ~ (1 | decade ) + (1 | region ) + inflation_annual_cpi + inflation_crises,
             data=pre.df,
             family=binomial,
             control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(currency_model_best_reg.dec)
```

